Introduction: a pipeline to analyse time-course gene expression data

### 2019-04-23 scripts/00_introduction.Rmd
FIX footnote

Gene expression studies provide simultaneous quantification of the level of
mRNA from all genes in a sample. High-throughput studies of gene expression
have a long history, starting with microarray technologies in the 1990s
through to single-cell technologies. While many expression studies are
designed to compare the gene expression in distinct groups, there is also a
long history of time-course expression studies, where the the gene expression
is compared across time by measuring mRNA levels from different samples across
time.\footnote{Because the collection of the mRNA is often destructive,
samples at different time points are generally from different biological
samples; longitudinal studies, for example tracking the same subject over
time, are certainly possible in certain settings, but not directly considered
here.} Such time course studies can vary from measuring a few distinct time
points, to sampling ten to twenty time points. Many longer time series are
particularly interested in investigating development over time. More recently,
single-cell studies track single cells through their development, and a single
cell is measured at a particular moment in its developmental progression -- a
value that is not know but estimated from the data as its "pseudo-time."

>>>>

Gene expression studies provide simultaneous quantification of the level of
mRNA from all genes in a sample. High-throughput studies of gene expression
have a long history, starting with microarray technologies in the 1990s
through to single-cell technologies. While many expression studies are
designed to compare the gene expression in distinct groups, there is also a
long history of time-course expression studies, where the the gene expression
is compared across time by measuring mRNA levels from different samples across
time [^1]. Such time course studies can vary from measuring a few distinct time
points, to sampling ten to twenty time points. Many longer time series are
particularly interested in investigating development over time. More recently,
single-cell studies track single cells through their development, and a single
cell is measured at a particular moment in its developmental progression -- a
value that is not know but estimated from the data as its "pseudo-time."

[^1]: Because the collection of the mRNA is often destructive, samples at
different time points are generally from different biological samples;
longitudinal studies, for example tracking the same subject over time, are
certainly possible in certain settings, but not directly considered here.
